\chapter{Introduction to the MetaCoq Project}
% \section{Introduction to the MetaCoq Project}

\section{The MetaCoq Project}
% \subsection{The MetaCoq Project}
MetaCoq is a project to formalize the core calculus, PCUIC, in Coq, and become
a platform to write tools that can manipulate Coq terms. The effort was complete
for a large part of the core language of Coq, with a few missing pieces:

\begin{itemize}
    \item Eta 
    \item Template Polymorphism
    \item SProps
    \item Modules
\end{itemize}

I will be tackling the last.

\section{Structure of the MetaCoq Project} The MetaCoq project is an
ambitious project aiming to provide a verified implementation of Coq, and for
its size, it is reasonably split into a few main components. From the layer
closest to the Coq language to the machine code, we have: TemplateCoq (Section
\ref{sec:mc-template}), PCUIC (Section \ref{sec:mc-pcuic}), followed by Safe
Checker, Erasure and beyond(Section \ref{sec:mc-beyond}).

Let us remind ourselves of the task of MetaCoq: we would like to see that the
OCaml representation of Coq is indeed correct and preserves the desired properties 
of the underlying theory. Since Coq has added many bells and whistles for its 
users, the terms of Coq definitely is much more complex than its underlying,
platonic type theoretical form. Therfore, MetaCoq has several stages for a term
to go through, stripping down to the bare minimum through the following few 
stages.

\subsection{TemplateCoq}
\label{sec:mc-template}

TemplateCoq is a quoting library for Coq: a Coq program that takes a Coq term,
and constructs an inductive data type that correspond to its kernel
representation in the OCaml implementation. This is the first layer of the
stripping of a Coq term, where the structures are preserved properly.

% TODO: Insert example

This allows one to turn a Coq program into a Coq internal representation along
with its associated environment structures, such as the definitions and
declarations in the environment.

\subsection{PCUIC}
\label{sec:mc-pcuic}

PCUIC is the Polymorphic Cumulative Calculus of Inductive Constructions. It is a
"cleaned up version of the term language of Coq and its associated type system,
shown equivalent to the one in Coq." (From MetqCoq website). In other words, it 
is a type theory that is as powerful as Coq can express, having the good
properties such as weakening, confluence, principality (that every term has a
principal type) etc.

A term generated in TemplateCoq can be converted into PCUIC term via a verified
process. The theory of PCUIC is then proven to have all the nice properties in
Coq.


\subsection{Safe Checker, Erasure and Beyond}
\label{sec:mc-beyond}

The core semantic operation of type theories are the reductions. The safechecker
is a verified "reduction machine, conversion checker and type checker" for PCUIC
terms. So far, we have the tools to start with a Coq term, first quoting into
TemplateCoq, then converted into a PCUIC term, and eventually has its type
checked in the Safe Checker via a fully verified process. As far as correctness
is concerned, this has already formed a verified end-to-end process of Coq's 
correctness.

The MetaCoq has further provided a verified Type and Proof erasure process from
PCUIC to untyped Lambda Calculus. The equivalence of this erased language is
can be evaluated in \emph{C-light} semantics, the subset of C accepted by
the CompCert verified compiler, which completes a maximally safe evaluation
toolchain for the language of Coq \sidecite{coqcoqcorrect}.
