\section{Motivation}
\label{sec:motivation}

\begin{frame}
    \frametitle{Motivation}
    \begin{itemize}
    \item Coq is a proof assistant based on Type Theory. In particular, the
        \textbf{P}olymorphic, \textbf{Cu}mulative Calculus of
        (co-)\textbf{I}nductive \textbf{C}onstructions (PCUIC).
    \item We trust the theory to be correct: PCUIC is proven strongly
        normalizing, reduction and conversion  are shown to be
        decidable\dots\pause
    \item But when we use Coq, we use the OCaml implementation. \pause
    \item Is this implementation also reliable? Correct?
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Bugs}
\begin{itemize}
\item On average, one critical bug a year!
\item Critical meaning possibly threatening the correctness of proved theorems -
    proving False in Coq!\pause
\item MetaCoq project in 2018 as a metaprogramming platform for Coq,
    building on existing systems such as TemplateCoq (2014).
\item Bonus: formalize the theory of Coq, PCUIC, using MetaCoq.\pause
\item Slogan: the Coq implementation you run every week/month.
\end{itemize}
\end{frame}

\begin{frame}
    \frametitle{What is the theory of Coq?}

    The type theory behind the Coq proof assistant is PCUIC:
    \begin{itemize}
    \item (Universe) \textbf{P}olymorphic 
    \item \textbf{Cu}mulative
    \item Calculus of (co-)\textbf{I}nductive \textbf{C}onstructions (CoC with Inductive Types)
    \item Calculus of Constructions (Coquand and Huet, 1988)
    \end{itemize}

\end{frame}

\begin{frame}
    \frametitle{MetaCoq - An Overview}

    \begin{itemize}
    \item From high level to low level:
    \item Coq - PCUIC - Machine Code \pause
    \item Coq - \textbf{TemplateCoq} - PCUIC - Machine Code \pause
    \item Coq - TemplateCoq - PCUIC - \textbf{Checker \& Erasure} - Machine Code\pause
    \item Coq Coq Correct! Verification of Type Checking and Erasure
        for Coq, in Coq (2020)\pause
    \item Caveats:\pause
    \begin{enumerate}
    \item Eta-expansion
    \item Template Polymorphism
    \item SProps (Proof-irrelevant Propositions)
    \item \textbf{Modules}
    \end{enumerate}
    
    \end{itemize}
\end{frame}


    
