\chapter{Project Plan for The Next Semester}
% \section{Project Plan for The Next Semester}

Since this is a joint thesis with the Mathematics department, also due to 
personal interest, for the remaining semester for this thesis, I am planning to:

\section{Implement Plain Coq Modules}
There are a few steps to this implementation that I can forsee:
\begin{enumerate}
\item Study the structure and conventions of the MetaCoq project. Since this is 
    a huge project, many common theorems have been abstracted into multiple
    functors which can generate a theorem given a corresponding object, such as
    % TODO: add functor example
\item Add the data structure of Plain Modules into the global environment which
    is quoted by TemplateCoq. This requires the modification of the current data
    structure containing declarations (minus modules). 
\item Ensure the environment remains well-defined with the current typing rules
    and correctness assertions. This requires the modifications to the existing 
    proofs on the environment (well-formedness, well-typedness).
\item Extend the TemplateCoq plugin (in OCaml) which transforms a Coq term into 
    a TemplateCoq representation, to translate a Coq module into the internal
    representation of Modules I implemented above. This requires OCaml knowledge
    specifically in the area of Coq plugins.
\item Now that syntatical transformation from Coq to TemplateCoq is complete, I
    need to define how these modules that live in the Global Environment will 
    be used. In particular, I need to implement the "canonical name" model as
    implemented in the Coq kernel.
\item Finally, the translation from TemplateCoq to PCUIC. In PCUIC,
    modules will cease to exist and references to definitions in modules will be
    treated as direct references to plain definitions outside modules. This is
    done by referring module paths to their canonical paths.
\item The last step in this project is to verify that the translation from
    TemplateCoq to PCUIC (and vice versa) is correct; that is, a PCUIC term is 
    well-formed and well-typed iff its translation is well-formed and well-typed
    in TemplateCoq.
\end{enumerate}

\section{Study Type Theoretic Results}

The modern field of Type Theory and Proof Assistants grow very quickly, with
many new fields of logic (modal logic, linear logic, seperation logic) etc and
their corresponding Type Theory being studied and made into proof assistants to 
verify results in these fields. Since there were not many relevant courses in my 
undergraduate curriculum, I decide to continue to study some classical results
alongside with the formalization project above.

The direction which I have chosen is towards a result important for the
implementation of proof assistants: the decidability of conversion
(correspondingly proof-checking) in a simpler type system, such as the Martin-Lof
Type Theory. So far, I have covered results from Simply Typed Lambda Calculus
including confluence and weak normalization, and its counterparts in proof theory,
namely sequent calculus, natural deduction as a normal form, and cut-elimination.